---
title: "Air Quality In National Park"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: https://github.com/linarettig/r_finalproject
    theme: simplex
---

```{r setup, include=FALSE}

library(shiny)
library(plotly)
library(leaflet)
library(tidyverse)
library(sf)

# Read in Data

```

Sidebar {.sidebar}
======================================================================


### **About the Data**

This dashboard uses data from the **TidyTuesday National Parks Visitation dataset** (2019-09-17 release), which contains annual visitor counts to U.S. National Parks from 1904–2016. Additional data on state populations and entrance fees are included to enable regional and temporal comparisons.

### **Purpose of the Dashboard**

The dashboard allows users to explore:
- Total annual visitation trends over time  
- Differences in visitation by U.S. region  
- Park-level visitation trends across different eras  
- Interactive views of individual park visitation patterns  

### **How to Navigate**

Use:
- The **sidebar controls** to filter years or regions  
- The **interactive plots** (hover to see values)  
- Tabs and rows to explore different panels in the dashboard  


```{r}

```


Main
======================================================================

Row
-----------------------------------------------------------------------


```{r}

```








