---
title: "Air Quality In National Park"
output:
  flexdashboard::flex_dashboard:
    social: menu
    source_code: https://github.com/linarettig/r_finalproject
runtime: shiny 
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(plotly)
library(leaflet)
library(tidyverse)
library(sf)
library(dplyr)
library(scales)


# Read in Data

locations <- read_csv("data/locations.csv")
visits <- read_csv("data/All National Parks Visitation 1904-2016.csv")
state_pop <- read_csv("data/state_pop.csv")
gas <- read_csv("data/gas_price.csv")
national_parks <- read_csv("data/national_parks.csv")

# Clean data and merge with state populations
visits_clean <- visits %>% 
  rename(visitors = Visitors,
         year = YearRaw,
         state = State,
         gnis_id = `Gnis Id`,
         park_name = `Unit Name`) %>% 
  mutate(year = as.numeric(as.character(year)),
         gnis_id = as.character(gnis_id)) %>% 
  group_by(gnis_id, state, year, park_name) %>% 
  summarize(visitors = sum(visitors, na.rm = TRUE), .groups = "drop") %>% 
  filter(!is.na(gnis_id), !is.na(year))

park_data <- visits_clean %>% 
  left_join(gas, by = "year") %>% 
  rename(gas_price = gas_current) %>% 
  select(-gas_constant)
  
park_data <- park_data %>% 
  left_join(state_pop, by = c("state" = "state", "year" = "year")) %>% 
  rename(state_population = pop)

locations_clean <- locations %>% 
  mutate(gnis_id = as.character(gnis_id))

park_data <- locations_clean %>% 
  right_join(park_data, by = "gnis_id") %>% 
  filter(!is.na(lon),
         !is.na(lat),
         !is.na(state_population),
         !is.na(gas_price)) %>% 
  mutate(lon = as.numeric(lon), lat = as.numeric(lat)) %>% 
  select(lon, lat, park_name, state, year, visitors, state_population, gas_price)

years_available <- sort(as.numeric(unique(park_data$year)))
max_visitors <- max(park_data$visitors, na.rm = TRUE)


# Ely Cleaned

park_visits <- read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2019/2019-09-17/national_parks.csv",
  show_col_types = FALSE
)

# Keep only National Parks with non-missing visitors
park_visits <- park_visits %>% 
  filter(unit_type == "National Park") %>% 
  filter(!is.na(visitors)) %>% 
  mutate(year_num = as.numeric(year))   

# IM region, years >= 2000 (no total rows)
im_parks <- park_visits %>% 
  filter(region == "IM",
         !is.na(year_num))

#grand total
im_total <- im_parks %>% 
  group_by(year_num) %>% 
  summarise(visitors = sum(visitors), .groups = "drop") %>% 
  mutate(unit_name = "Total Visitation")

im_parks <- bind_rows(im_parks, im_total)

# Vector of IM park names for the dropdown
im_park_list <- sort(unique(im_parks$unit_name))

```



All National Parks
======================================================================


### National Parks Desriptions 

```{r Year slider and Dropdown}
sliderInput("year_select", "Select Year:",
            min = min(years_available),
            max = max(years_available),
            value = 1929,
            step = 1,
            sep = "",
            animate = animationOptions(interval = 1000, loop = FALSE))
```

```{r}
renderText({
  paste("Rows of filtered data for year", input$year_select, ":", nrow(filtered_data()))
})
```



### **About the Data**

This dashboard uses data from the **TidyTuesday National Parks Visitation dataset** (2019-09-17 release), which contains annual visitor counts to U.S. National Parks from 1904–2016. Additional data on state populations and entrance fees are included to enable regional and temporal comparisons.

### **Purpose of the Dashboard**

The dashboard allows users to explore:
- Total annual visitation trends over time  
- Differences in visitation by U.S. region  
- Park-level visitation trends across different eras  
- Interactive views of individual park visitation patterns  

### **How to Navigate**

Use:
- The **sidebar controls** to filter years or regions  
- The **interactive plots** (hover to see values)  
- Tabs and rows to explore different panels in the dashboard 




### Pane 1-  Park Visitor


Row 2
------------------------------------------


### Pane 2- Interactive Visitation Trend for Denali National Park
```{r interactive_plot, echo=TRUE}

library(plotly)
unique(park_visits$unit_name)

str(park_visits$year)

unique(park_visits$year)

demo <- park_visits %>% filter(str_detect(unit_name, "Denali"))

summary(demo$year)

demo <- park_visits %>% 
  filter(str_detect(unit_name, "Denali"))

p <- ggplot(demo, aes(x = year, y = visitors)) +
     geom_line(color = "steelblue", linewidth = 1) +
     geom_point(size = 0.7, alpha = 0.6) +
     scale_x_continuous(breaks = seq(min(demo$year), max(demo$year), by = 10)) +
     labs(title = "Visitation Trend: Denali National Park",
          x = "Year",
          y = "Visitors")

ggplotly(p)

```

Row 
=======
Column {data-width=700}

-----------------------------------------------------------------------

### Pane 3- Park Visitors, State Pop and Gas Prices by year

```{r pane3_leaflet, echo=TRUE}
# Reactive expression to filter data based on slider input 
filtered_data <- reactive({
  req(input$year_select)
  park_data %>% 
    filter(year == input$year_select) %>% 
    distinct(lon, lat, .keep_all = TRUE)
})

leafletOutput("leaflet_map")
```


```{r}
output$leaflet_map <- renderLeaflet({
  leaflet() %>% 
    addTiles() %>% 
    setView(lng = -98.5, lat = 39.8, zoom = 4)
})

observe({
  data_to_plot <- filtered_data()
  
  data_to_plot <- data_to_plot %>% 
    mutate(
      radius = sqrt(visitors) / sqrt(max_visitors) * 20 + 3,
      popup_html = paste(
        "<b>", park_name, "</b> (", state, ")", "<br>",
        "Year: ", year, "<br>",
        "Visitors: ", formatC(visitors, format="f", big.mark=",", digits=0), "<br>",
        "State Pop: ", formatC(state_population, format="f", big.mark=",", digits=0), "<br>",
        "Avg Gas Price (Current $): $", round(gas_price, 2)
      )
    )
  leafletProxy("leaflet_map", data = data_to_plot) %>% 
    clearMarkers() %>% 
    addCircleMarkers(
      lng = ~lon,
      lat = ~lat,
      radius = ~radius,
      fillColor = "#1f78b4", 
      color = "#000000",
      weight = 0.5, 
      opacity = 1, 
      fillOpacity = 0.8, 
      popup = ~popup_html
    )
})
```






InterMountain Region 
================================================================================

Row 
-------------------------------------------------------------------------------

### Visitors by year

```{r}
selectInput(
  inputId  = "park_choice",
  label    = "Choose a park:",
  choices  = im_park_list,
  selected = "Yellowstone National Park"
)

renderPlot({
  req(input$park_choice)

  df <- im_parks %>% 
    filter(unit_name == input$park_choice)

  ggplot(df, aes(x = year_num, y = visitors)) +
    geom_line(linewidth = 1.2) +
    geom_point() +
    labs(
      title = paste("Annual visitation –", input$park_choice),
      x = "Year",
      y = "Number of visitors"
    ) +
    theme_minimal(base_size = 14)
})
```

## AirQuality Starts HERE - Eliud


<<<<<<< HEAD
=======

>>>>>>> 43f322a79934d9837d3c146d0ff10eb09b3ef4f2
```{r}
#slider for AQI 
air <- read_csv("data/Air quality.csv")

air_im <- air %>% 
  filter(State %in% c("Montana", "Wyoming", "Utah", "Colorado","Arizona", "New Mexico", "Texas", "Oklahoma"))
```


```{r}
#slider 


#Air quality table 


```
<<<<<<< HEAD



=======
>>>>>>> 43f322a79934d9837d3c146d0ff10eb09b3ef4f2

