---
title: "Air Quality In National Park"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: https://github.com/linarettig/r_finalproject
    theme: simplex
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(plotly)
library(leaflet)
library(tidyverse)
library(sf)
library(dplyr)
library(scales)


# Read in Data

locations <- read_csv("data/locations.csv")
visits <- read_csv("data/All National Parks Visitation 1904-2016.csv")
state_pop <- read_csv("data/state_pop.csv")
gas <- read_csv("data/gas_price.csv")
national_parks <- read_csv("data/national_parks.csv")

# Clean data and merge with state populations
vis_data_pop <- visits %>% 
  rename(visitors = Visitors,
         year = YearRaw,
         state = State,
         gnis_id = `Gnis Id`,
         park_name_display = `Unit Name`) %>% 
  mutate(year = as.numeric(as.character(year)),
         gnis_id = as.character(gnis_id)) %>% 
  group_by(gnis_id, state, year, park_name_display) %>% 
  summarize(visitors = sum(visitors, na.rm = TRUE), .group = "drop") %>% 
  left_join(state_pop, by = c("state" = "state", "year" = "year")) %>% 
  rename(state_population = pop) %>% 
  filter(!is.na(year), !is.na(state_population))

# Merge with Location
park_year_data <- locations %>% 
  mutate(gnis_id = as.character(gnis_id)) %>% 
  rename(lon = lon, lat = lat) %>% 
  right_join(vis_data_pop, by = "gnis_id") %>% 
  filter(!is.na(lon), !is.na(lat)) %>% 
  select(lon, lat, state, park_name_display, year, visitors, state_population)


# Create set of available years
years_available <- sort(as.numeric(unique(park_year_data$year)))

# Calculate max visitation for marker scaling 
max_visitors <- max(park_year_data$visitors, na.rm = TRUE)

```


```{r}
# Read in Data
park_visits <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-09-17/national_parks.csv")

# Keep only National Parks
park_visits <- park_visits %>% filter(unit_type == "National Park")

# Create year categories
park_visits <- park_visits %>%
  mutate(
    year_cat = case_when(
      year <= 1950 ~ "1904–1950",
      year <= 1980 ~ "1951–1980",
      TRUE ~ "1981–2016"
    ),
    year_cat = factor(year_cat)
  )

# Create summarized dataset for Plot 1
park_summary <- park_visits %>%
  group_by(region, year_cat, unit_name) %>%
  summarise(total_visitors = sum(visitors, na.rm = TRUE), .groups = "drop")

```








Sidebar {.sidebar}
======================================================================


)
```{r Year slider}
sliderInput("year_select", "Select Year:",
            min = min(years_available),
            max = max(years_available),
            value = min(years_available),
            step = 1,
            sep = "",
            animate = animationOptions(interval = 1000, loop = FALSE))
```


### **About the Data**

This dashboard uses data from the **TidyTuesday National Parks Visitation dataset** (2019-09-17 release), which contains annual visitor counts to U.S. National Parks from 1904–2016. Additional data on state populations and entrance fees are included to enable regional and temporal comparisons.

### **Purpose of the Dashboard**

The dashboard allows users to explore:
- Total annual visitation trends over time  
- Differences in visitation by U.S. region  
- Park-level visitation trends across different eras  
- Interactive views of individual park visitation patterns  

### **How to Navigate**

Use:
- The **sidebar controls** to filter years or regions  
- The **interactive plots** (hover to see values)  
- Tabs and rows to explore different panels in the dashboard 




```{r}
# Where info and slide and drop down will be 
```


Main
======================================================================

Row
-----------------------------------------------------------------------

### Park Visitor

```{r echo=TRUE}
park_summary %>%
  ggplot(aes(x = total_visitors, y = fct_reorder(unit_name, total_visitors), color = year_cat)) +
  geom_point(size = 2, alpha = 0.7) + facet_wrap(~ region, scales = "free_y") +
  labs(title = "Total Visitors to U.S. National Parks by Region and Era",
       x = "Total Annual Visitors", y = "National Park", color = "Era") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 6))


```

Row 
-----------------------------------------------------------------------

### Park Visitors, State Pop and Gas Prices by year

```{r Leaflet map}
# Reactive expression to filter data based on slider input 
filtered_data <- reactive({
  req(input$year_select)
  park_year_data %>% 
    filter(year == input$year_select) %>% 
    distinct(lon, lat, .keep_all = TRUE)
})

# Leaflet map base layer 
renderLeaflet({
  leaflet() %>% 
    addTiles() %>% 
    setView(lng = -98.5, lat = 39.8, zoom = 3)
})

observe({
  data_to_plot <- filtered_data()
  
  data_to_plot <- data_to_plot %>% 
    mutate(
      radius = sqrt(visitors) / sqrt(max_visitors) * 20 + 3,
      popup_text = paste(
        "<b>", park_name_display, "</b><br>",
        "Year: ", year, "<br>",
        "Visitors: ", formatC(visitors, format="f", big.mark=",", digits=0), "<br>",
        "State: ", state, "<br>",
        "State Population: ", formatC(state_population, format="f", big.mark=",", digits=0)
      )
    )
  leafletProxy("Leafket map", data = data_to_plot) %>% 
    clearMarkers() %>% 
    addCircleMarkers(
      lng = ~lon,
      lat = ~lat,
      radius = ~radius,
      fillColor = "#0072B2", 
      color = "#000000",
      weight = 0.5, 
      opacity = 1, 
      fillOpacity = 0.8, 
      popup = ~popup_text
    )
})
```



